### 2020-01-05 grand_challenge/astro_grand_challenge.tex
some feedback

While a variety of approximations are used in astrophysics to model
hydrodynamic flows, for stellar explosions, conservative finite-volume
methods are the most popular.  Adaptive mesh refinement (AMR) allows
us to increase the range of lengthscales captures.  Our group has been
developing the AMReX Astrophysics Suite of open-source AMR
multiphysics hydrodynamics
codes\footnote{\url{https://github.com/AMReX-Astro/}} for over a
decade.  Two codes in particular address stellar explosions: MAESTROeX
is a low Mach number hydrodynamics code for modeling the highly
subsonic stellar convection leading up to explosion and Castro is a
fully compressible radiation hydrodynamics code that efficiently
models the explosive phase.  These codes share a common framework
(adaptive mesh library, microphysics routines).

@@

To highlight the way that new computer architectures enable new
science, we focus on our current work on XRBs.  Our interest is in
modeling the nucleosynthesis as a burning front propagates laterally
through the accreted fuel layer on a neutron star.  Due to the range
of length and timescales, various approximations have been used in the
literature.  Our group started performing two-dimensional
hydrodynamical simulations of the flame propagating using the NERSC
Cori and OLCF Titan machines over a year ago.  These calculations were
the first to perform full hydrodynamics in both the lateral and
vertical directions, allowing us to understand the complex dynamics of
the flame front with realistic microphysics.  The simulations were
challenging, and we needed to make approximations to both the length
scales (using artificially high neutron star rotation rates to compress
the lateral scales) and time scales (boosting the flame speed).
Nevertheless, we were able to complete the first resolved calculations
of hydrodynamical flames in X-ray bursts.  Taking advantage of GPUs,
we are now able to eliminate these approximations in our current runs on the
OLCF Summit machine.  The same code is used,
but running entirely on the GPUs, it is 5--10$\times$ faster than
CPUs.  We are on target to perform the first three-dimensional
simulations this year as well as to explore the more interesting case
of mixed H/He bursts, which involve more complicated nuclear
reactions.

@@

New architectures bring new programming challenges, and to date we've
managed to keep pace with these developments.  A key point is that
funding needs to be provided to aid in preparing codes, not just for
the science, especially for novel architectures.  NSF should recognize
this need.  To ensure the largest benefit to the community,
application codes for these machines should be Open Source to the
fullest extend allowed.

>>>>

While a variety of approximations are used in astrophysics to model
hydrodynamic flows, for stellar explosions, conservative finite-volume
methods are the most popular.  Adaptive mesh refinement (AMR) allows
us to increase the range of lengthscales captures.  Our group has been
developing the AMReX Astrophysics Suite of open-source AMR
multiphysics hydrodynamics
codes\footnote{\url{https://github.com/AMReX-Astro/}} for over a
decade.  Two codes in particular address stellar explosions: MAESTROeX
is a low Mach number hydrodynamics code for modeling the highly
subsonic stellar convection leading up to explosion and Castro is a
fully compressible radiation hydrodynamics code that efficiently
models the explosive phase.  These codes share a common framework
(adaptive mesh library, microphysics routines), are written in C++ and Fortran,
and use MPI+OpenMP or MPI+CUDA.

@@

To highlight the way that new computer architectures enable new
science, we focus on our current work on XRBs.  Our interest is in
modeling the nucleosynthesis as a burning front propagates laterally
through the accreted fuel layer on a neutron star.  Due to the range
of length and timescales, various approximations have been used in the
literature.  Our group started performing two-dimensional
hydrodynamical simulations of the flame propagating using the NERSC
Cori and OLCF Titan (on CPUs)  machines over a year ago.  These calculations were
the first to perform full hydrodynamics in both the lateral and
vertical directions, allowing us to understand the complex dynamics of
the flame front with realistic microphysics.  The simulations were
challenging, and we needed to make approximations to both the length
scales (using artificially high neutron star rotation rates to compress
the lateral scales) and time scales (boosting the flame speed).
Nevertheless, we were able to complete the first resolved calculations
of hydrodynamical flames in X-ray bursts.  Taking advantage of GPUs,
we are now able to eliminate these approximations in our current runs on the
OLCF Summit machine.  The same code is used,
but running entirely on the GPUs, it is 5--10$\times$ faster than
CPUs.  We are on target to perform the first three-dimensional
simulations this year as well as to explore the more interesting case
of mixed H/He bursts, which involve more complicated nuclear
reactions.

@@

New architectures bring new programming challenges, and to date we've
managed to keep pace with these developments.  A key point is that
funding needs to be provided to aid in preparing codes, not just for
the science, especially for novel architectures.  NSF should recognize
this need.  The software stack on the next generation machine is as
important as the hardware.  It should support the popular
community-supported programming models today and have wide support for
C/C++ and Fortran.  To ensure the largest benefit to the community,
application codes for these machines should be Open Source to the
fullest extend allowed.
